Equalizer 2.0\\Convergence of a Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
Improved results wording
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@47 db7c848b-7755-4df2-b763-8796797afa52

A critical piece for data distribution performance are the characteristics of
the data compression algorithms. This microbenchmark compresses a set of binary
files to precalculate the speed and compression ratio of the various
engines. \fig{fCompressor} shows the compression and decompression speed in
gigabyte per second as well as the size of the compressed data. The different
ZSTD$x$ engines use the ZStandard compression library at level $x$. The
measurements were performed on a single, isolated node.

@@

RLE compression has a very low overhead, but does not compress the data will
since it only removes ``blank space'' in the data. The Snappy compression, used
as default in \textsf{Collage}, achieves the same compression ratio as the LZ variants at
a much higher speed. The ZStandard compressor has roughly the same speed as the
LZ variants at the lowest compression level, but provides significantly better
compression. At higher compression levels it can improve the compression ratio
slightly, but at a high cost for the compression speed.

The compression ratio for the models used in the following benchmark deviate
from this averaged distribution. \fig{fCompressorDetail} shows the compression
ratios for the triangle and volume data.

@@

The ply data is little compressible, the default compressor achieves a 10\%
reduction. This is due to a high entropy in the data and the dominant use of
floating point representations. Overall the profile is similar to the generic
benchmark, at a smaller compression rate.

The volume data on the other hand is sparsely populated and using integer (byte)
representation, which is easier to compress. The naive RLE implementation
already achieves a good compression, showing that the smaller volume contains at
most 28\% empty space and the bigger volume at most 90\%. Snappy and ZStandard
can reduce the data much further, reducing the data to a few megabytes.

@@

In this section we present how data distribution and synchronization performs in
real-world applications. We extracted the existing data distribution code in a
PLY and a volume renderer into a benchmark application to measure the time to
initially map all the objects on the render client nodes, and to perform a
commit-sync of the full data set after mapping has been established. All figures
observe a noticable measurement jitter due to other services running on the
cluster. The details of the benchmark algorithm can be found in the
implementation~\cite{eqObjectBench}. We used the same data sets as in the
previous section. After $8$ processes, a nodes start to run two processes per
node, which share the CPU and network interface.

Object mapping is measured using the following settings: \textbf{none}
distributes the raw, uncompressed, and unbuffered data, \textbf{compression}
uses the Snappy compresser to distribute unbuffered data, \textbf{buffered}
reuses uncompressed, serialized data for mappings of other nodes, and
\textbf{compression buffered} compresses this buffer. Unbuffered operations need
to reserialize, and potentially recompress, the master object data for each
slave node. Each slave instance needs to deserialize and decompress the data.

@@

The spike volume data set shown has a significant compression ratio, which
reflects in the results in \fig{fSpikesDist}. Futhermore, only a single
distributed object is used to transmit the volume information in one chunk.
Compression for this data is benefitial for transmitting data over a 10Gb/s link
even for a single slave process only (970 MB/s vs 570 MB/s raw data transfer).
Buffering has little benefit since the serialization of volume data is trivial.
Buffered compression makes a huge difference, since the compression cost can be
amortized over $n$ nodes, reaching raw data transmission rates of 3.7 GB/s with
the default Snappy compressor and at best 4.4 GB/s with ZStandard at the lowest
compression ratio.

@@

\fig{fBeechnutDist}

@@

Distribute large ply and raw model to 1..16 nodes using TCP, IB and RSP

>>>>

A critical piece for data distribution performance are the characteristics of
the data compression algorithm. Our microbenchmark compresses a set of binary
files, precalculating the speed and compression ratio of the various engines.
\fig{fCompressor} shows the compression and decompression speed in gigabyte per
second as well as the size of the compressed data relative to the uncompressed
data. The different ZSTD$x$ engines use the ZStandard compression library at
compression level $x$. The measurements were performed on a single, isolated
node.

@@

RLE compression has a very low overhead, but does not compress the data, since
it only removes ``blank space'' in the data. The Snappy compression, used as
default in \textsf{Collage}, achieves the same compression ratio as the LZ
variants at a much higher speed. The ZStandard compressor has roughly the same
speed as the LZ variants at the lowest compression level, but provides
significantly better compression. At higher compression levels it can improve
the compression ratio slightly, but at a high cost for the compression speed.

The compression ratio for the models used in the following section deviate from
this averaged distribution. \fig{fCompressorDetail} shows the compression ratios
for the triangle and volume data.

@@

The ply data is little compressible, the default compressor achieves a 10\%
reduction. This is due to a high entropy in the data and the dominant use of
floating point values. Overall the profile is similar to the generic benchmark,
at a smaller compression rate.

The volume data on the other hand is sparsely populated and using integer (byte
and short) values, which is easier to compress. The naive RLE implementation
already achieves a good compression, showing that the smaller volume contains at
most 28\% empty space and the bigger volume at most 43\%. Snappy and ZStandard
can reduce the spike data much further, reducing the data to a few megabytes.
Surprisingly, the beechnut data set does not yield significant higher
compression with the modern Snappy and ZStandard libraries.

@@

In this section we present how data distribution and synchronization performs in
real-world applications. We extracted the existing data distribution code in a
PLY (eqPly) and a volume renderer (Livre) into a benchmark application to
measure the time to initially map all the objects on the render client nodes,
and to perform a commit-sync of the full data set after mapping has been
established. All figures observe a noticable measurement jitter due to other
services running on the cluster. The details of the benchmark algorithm can be
found in the implementation~\cite{eqObjectBench}. We used the same data sets as
in the previous section. After eight processes, a nodes start to run two
processes per node, which share CPU, memory and network interface bandwidth.

Object mapping is measured using the following settings: \textbf{none}
distributes the raw, uncompressed, and unbuffered data, \textbf{compression}
uses the Snappy compressor to distribute unbuffered data, \textbf{buffered}
reuses uncompressed, serialized data for mappings of other nodes, and
\textbf{compression buffered} compresses this buffer. Unbuffered operations need
to reserialize, and potentially recompress, the master object data for each
slave node. Each slave instance needs to deserialize and decompress the data.

@@

The volume data sets are each distributed in a single object, serializing the
raw volume buffer. The spike volume data set has a significant compression
ratio, which reflects in the results in \fig{fSpikesDist}. Compression for this
data is benefitial for transmitting data over a 10~Gb/s link even for a single
slave process only (970 MB/s vs 570 MB/s raw data transfer). Buffering has
little benefit since the serialization of volume data is trivial. Buffered
compression makes a huge difference, since the compression cost can be amortized
over $n$ nodes, reaching raw data transmission rates of 3.7 GB/s with the
default Snappy compressor and at best 4.4 GB/s with ZStandard at the lowest
compression ratio.

@@

The distribution of the beechnut data set also behaves as expected
(\fig{fBeechnutDist}). Due to the larger object size, uncompressed transmission
is slightly faster compared to the spike data set at 700 MB/s, and compressed
transmission does not improve the mapping performance, likely due to increased
memory pressure caused by the large data size. The comparison of the various
compression engines is consisten with the benchmarks in
\fig{fCompressorDetail}; RLE, Snappy and the LZ variants are very close to each
other, and ZSTD1 can provide better performance after four nodes due to the
better compression ratio.

@@

\FIXME{Missing: david on 10GE, spike comparisoon 10GE-IB-RSP}
